---
title: "Model fit showcase"
author: "Gilles Lijnzaad"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
    number_sections: TRUE
    highlight: tango
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(purl = knitr::hook_purl) # creates an R file upon knitting the Rmd
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 8)
```

> In this document, I explain the code for my standard Stan model of the climate-RL task (without confirmation bias) and demonstrate parameter recovery.

# Helper files {.unnumbered}
In this R markdown I use code from the following files: [sim_utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#sim_utilsR), [plot_utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#plot_utilsR), [fit_utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#fit_utilsR), and [utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#utilsR).

# Creating the data

--- 

## Running the simulation
```{r create-data, comment = NA}
rm(list = ls())
setwd("~/research/climate-RL-mod/2_modeling_showcase")
main_dir <- "~/research/climate-RL-mod/"
util_dir <- paste0(main_dir, "9_utilities/")
mod_dir <- paste0(main_dir, "0_models/")
model_path <- paste0(mod_dir, "1_std.stan")

sim <- new.env()
source(paste0(mod_dir, "1_std.R"), local = sim) # access functions using sim$fun()

params <- list(
  n_part = 50,
  n_trials = 30,
  LR_group = 0.4,
  inv_temp_group = 0.5,
  initQ_dev_group = 2,
  mu_R_group = 5.5,
  sigma_R_group = 2
)

sim_dat <- sim$run(params)

cat(paste0("PARAMETER SETTINGS:"), capture.output(dplyr::glimpse(params)), sep = "\n")
cat(paste0("SIMULATED DATA:"), capture.output(dplyr::glimpse(sim_dat)), sep = "\n")
```

## Inspecting the data
Just so I know what to expect of our model, I include plots of the simulated data: Q and choices over time.

```{r inspecting-data, fig.height = 4}
plot <- new.env()
source(paste0(util_dir, "plot_utils.R"), local = plot) # access functions using plot$fun()

plot$sim_plots(sim_dat, params)
```

# Explaining the model

--- 

I show the Stan model code bit by bit to explain what is happening. The full code can be found in `1_std.stan`.

## `data` block
The data we supply to the model consist of:

1. the number of participants: `n_part`;
1. the number of trials per participant: `n_trials`;
1. an array of choices that a participant made: `choice_c[n_part, n_trials]`, coded as (1) consistent with the very first choice (hence `choice_c`) or (2) not consistent with the first choice;
1. an array of ratings that a participant gave: `R[n_part, n_trials]`. These are analogous to the array `outcomes[n_trials]` in the `2arm_bandit_example` (see [Stan code](https://github.com/CCS-Lab/hBayesDM/blob/develop/commons/stan_files/bandit2arm_delta.stan) for this example);
1. the mean rating of a participant, as calculated from the data: `mu_R[n_part]`.

```{r explmod-data, comment = NA}
mod_code <- readLines(model_path)
start <- grep("data \\{", mod_code)[1]
end <- grep("transformed data \\{", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `transformed data` block
Currently unused.
```{r explmod-transf-data, eval = FALSE, comment = NA}
start <- grep("transformed data \\{", mod_code)
end <- grep("parameters \\{", mod_code)[1] - 2
cat(mod_code[start:end], sep = "\n")
```

## `parameters` block
In the current model, we have three free parameters:

1. learning rate `LR`;
1. inverse temperature `inv_temp`;
1. deviation from the mean for initial Q: `initQ_dev`. This is related to the actual initial Qs as: 
  > `initQ_pref = mu_R + initQ_dev`
  > `initQ_nonpref = mu_R - initQ_dev`

[explanation here]

```{r explmod-params, comment = NA}
start <- grep("parameters \\{", mod_code)[1]
end <- grep("transformed parameters \\{", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `transformed parameters` block
[explanation here]
```{r explmod-transf-params, eval = FALSE, comment = NA}
start <- grep("transformed parameters \\{", mod_code)
end <- grep("model \\{", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `model` block
### The priors
[explanation here]
```{r explmod-model-priors, comment = NA}
start <- grep("model \\{", mod_code)
end <- grep("participant loop", mod_code)[1] - 2
cat(mod_code[start:end], sep = "\n")
```

### Participant loop
[explanation here]
```{r explmod-model-part, comment = NA}
start <- grep("participant loop", mod_code) + 1
end <- grep("trial loop", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

### Trial loop
```{r explmod-model-trial, comment = NA}
start <- grep("trial loop", mod_code) + 1
end <- grep("generated quantities", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `generated quantities` block
[explanation here]
```{r explmod-gen-quant, comment = NA}
start <- grep("generated quantities", mod_code)
end <- length(mod_code)
cat(mod_code[start:end], sep = "\n")
```

# Fitting the model

--- 

## Single run: chosen parameters
### Do the run

```{r run-model}
dat_dir <- paste0(main_dir, "3_param_recovery/1_std/1_run/")
dat_file <- paste0(dat_dir, "sim_dat_1.json")

fitting <- new.env()
source(paste0(util_dir, "fit_utils.R"), local = fitting) # access functions using fitting$fun()
model <- cmdstan_model(model_path)
# draws <- fitting$fit(model, dat_file, return = "draws", show_iteration_progress = TRUE)
# saveRDS(draws, paste0(dat_dir, "draws.rds"))
```

### Group-level parameters
```{r posterior-plots}
free_params_group <- c("LR_group", "inv_temp_group", "initQ_dev_group")

draws <- readRDS(paste0(dat_dir, "draws.rds"))
draws <- draws %>%
  rename(setNames(paste0("means[", seq_along(free_params_group), "]"),
                  free_params_group))
to_plot <- free_params_group
plot$posterior_densities(draws, to_plot, params)
```
```{r posterior-table}
util <- new.env()
source(paste0(util_dir, "utils.R"), local = util) # access functions using util$fun()
to_inspect <- free_params_group
util$print_posterior_table(draws, params, to_inspect)
```

### Participant-level parameters
```{r sim-vs-fit}
participant_params <- rjson::fromJSON(file = paste0(dat_dir, "sim_param_settings_1.json"))
free_params_pp <- gsub("_group", "", free_params_group)
plot$pp_level_param_fit(draws, free_params_pp, participant_params)
```

TODO: for inv_temp it seems okay, but for the other parameters the participant-level estimations are very similar to the group-level estimations. Seems like a mistake?

## Many runs: random parameters
To check parameter recovery across a range of parameter values, I run the model a bunch of times with different parameter settings. More specifically, each parameter value is chosen randomly from its whole theoretical range.
```{r many-runs}
n_runs <- 100
dat_dir <- paste0(main_dir, "3_param_recovery/1_std/100_runs/")

sim$run_many(params, dat_dir, n_runs)
# running this bit below takes at least an hour
# fitting$fit_many(dat_dir, model_path, fit_data_type = "draws", dat_dir, n_runs)
```

```{r inspect-many-runs}
sim_params <- data.frame(k = 1:n_runs)
fit_params <- data.frame(k = 1:n_runs)

for (k in 1:n_runs) {
  sim_file <- paste0(dat_dir, "param_settings_", sprintf("%03d", k), ".json")
  sim_dat <- rjson::fromJSON(file = sim_file)

  fit_file <- paste0(dat_dir, "draws_", sprintf("%03d", k), ".rds")
  fit_dat <- readRDS(fit_file) %>%
    rename(setNames(paste0("means[", seq_along(free_params_group), "]"),
                    free_params_group))

  for (p in free_params_group) {
    sim_params[[p]][k] <- sim_dat[[p]]
    fit_params[[p]][k] <- median(fit_dat[[p]])
  }
}

plot$many_runs_param_fit(sim_params, fit_params, free_params_group)
```