---
title: "Stan models"
author: "Gilles Lijnzaad"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: TRUE
    number_sections: true
    highlight: tango
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(purl = knitr::hook_purl) # creates an R file upon knitting the Rmd
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 4)
```

# The data
## Loading them
The data are coming from the R simulation. The `sim_param_settings` file shows what parameters we ran the simulation with (parameters `LR`, `inv_temp`, `initQF`, `initQU`, `mu_R`, `sigma_R`), so we're sure what data we're looking at. The `sim_dat` file shows the simulated data (`T`, `choice`, `R`).

```{r loading-data, message=FALSE}
rm(list = ls())
sim_dir <- "~/research/climate-RL/R_simulation/"
library(rjson)
param_settings <- fromJSON(file = paste0(sim_dir, "sim_param_settings.json"))
data <- fromJSON(file = paste0(sim_dir, "sim_dat.json"))
library(dplyr)
glimpse(param_settings)
glimpse(data)
```

## Inspecting them
Just so we know what to expect of our model, we include plots of the Q values and choices over time. The plots are generated by `sim.Rmd`.

```{r inspecting-data, echo = FALSE}
mod_dir <- "~/research/climate-RL/stan_models/my_modeling/"
plot_Q <- readRDS(paste0(mod_dir, "plot_Q.rds"))
plot_choice <- readRDS(paste0(mod_dir, "plot_choice.rds"))

library(grid)
library(gridExtra)
grid.arrange(plot_Q, plot_choice, nrow = 1,
             top = textGrob("Simulated data", gp = gpar(fontsize = 20, font = 2)))

text <- paste0("LR = ", param_settings$LR,
              "\ninv_temp = ", param_settings$inv_temp,
              "\ninitQF = ", param_settings$initQF,
              "\ninitQU = ", param_settings$initQU,
              "\nmu_R_F = ", param_settings$mu_R_F,
              "\nmu_R_U = ", param_settings$mu_R_U,
              "\nsigma_R = ", param_settings$sigma_R)
grid.text(text, x = unit(0.98, "npc"), y = unit(0.87, "npc"), hjust = 1, vjust = 1)

```

# Defining the model
I show the Stan model code bit by bit to explain what is happening.

## `data` block
The data we supply to the model consist of:

1. the number of participants: `n_part`;
1. the number of trials per participant: `T`;
1. an array of choices that a participant made: `choice[n_part, T]`;
1. an array of ratings that a participant gave: `R[n_part, T]`. These are analogous to the array `outcomes[T]` in the `2arm_bandit_example` (see [Stan code](https://github.com/gilleslijnzaad/climate-RL-modeling/blob/main/stan_models/2arm_bandit_example/2arm_bandit_example.stan) for this example).

```{r defmod-data}
mod <- ""
mod <- paste0(mod, 
"data {
  int<lower=1> n_part;
  int<lower=1> T;
  array[n_part, T] int<lower=1, upper=2> choice;
  array[n_part, T] int<lower=0, upper=10> R;
  int<lower=0, upper=10> initQF;
  int<lower=0, upper=10> initQU;
}
")
```

## `parameters` block
In the current model, we have two free parameters:

1. learning rate 
1. inverse temperature

I copied the `2arm_bandit_example` ([see code](https://github.com/gilleslijnzaad/climate-RL-modeling/blob/main/stan_models/2arm_bandit_example/2arm_bandit_example.stan)) in defining "raw parameters" that we later convert using inverse logit transformation.

When I make the model hierarchical, these parameters will be group-level means and standard deviations.

```{r defmod-params}
mod <- paste0(mod, "
parameters {
  real<lower=0, upper=1> LR;
  real<lower=0, upper=5> inv_temp;
}
")
```

## `transformed parameters` block
This approach copies the `2arm_bandit` approach. 
```{r defmod-transf-params}
mod <- paste0(mod, "
// transformed parameters {
// }
")
```

## `model` block
### The priors
Again, following the `2arm_bandit_example` approach.
```{r defmod-model-priors}
mod <- paste0(mod, "
model {
  // priors: all uninformative
  LR ~ uniform(0, 1);
  inv_temp ~ uniform(0, 5);
")
```

### Participant loop
For every participant I initialize an array for holding the Q values. I also create a temporary vector `Q_t` that holds the two Q values (F, U) for the *current* trial, because otherwise the `categorical_logit` wouldn't work, somehow. This works for now, and maybe in the future I will find a more elegant solution.
```{r defmod-model-inits}
mod <- paste0(mod, "
  for (j in 1:n_part) {
    array[T, 2] real Q;
    Q[1, 1] = initQF;
    Q[1, 2] = initQU;
    vector[2] Q_t;

    real pred_err;
")
```

### Trial loop
```{r defmod-data-trials}
mod <- paste0(mod, "
    for (t in 1:T) {
      Q_t = to_vector(Q[t]);

      // sample choice (0 is F, 1 is U) via softmax
      choice[j, t] ~ categorical_logit(inv_temp * Q_t);

      // prediction error
      if (choice[j, t] == 0) {
        pred_err = R[j, t] - Q[t, 1];
      } else {
        pred_err = R[j, t] - Q[t, 2];
      }

      // update value (learn)
      if (t < T) {    // no updating in the very last trial
        if (choice[j, t] == 0) {
          Q[t+1, 1] = Q[t, 1] + LR * pred_err;
          Q[t+1, 2] = Q[t, 2];
        } else {
          Q[t+1, 1] = Q[t, 1];
          Q[t+1, 2] = Q[t, 2] + LR * pred_err;
        }
      }
    }
  }
}
")
```

## Save Stan code
```{r save-stan}
write(mod, file = paste0(mod_dir, "climate-RL.stan"))
```

# Fitting the model
## Run it
```{r run-model}
library(cmdstanr)

# function allows for easily changing whether you want to refit or use a saved fit
it <- 10000
fit_model <- function(refit) {
  if (refit) {
    m <- cmdstan_model(paste0(mod_dir, "climate-RL.stan"))
    data_file <- paste0(sim_dir, "sim_dat.json")
    fit <- m$sample(
      data = data_file,
      iter_sampling = it,
      chains = 1,
      thin = 1,
      iter_warmup = it / 2,
      refresh = it / 5,
      seed = 1234
    )
    fit$save_object(file = paste0(mod_dir, "climate-RL_fit.rds"))
  } else {
    fit <- readRDS(file = paste0(mod_dir, "climate-RL_fit.rds"))
  }
  return(fit)
}

fit <- fit_model(refit = TRUE)
```

## Inspect results
### Helper function: plotting
```{r}
library(ggplot2)
my_teal <- "#008080"
my_pink <- "#ff00dd"
my_blue <- "#00ccff"
my_colors <- c(my_teal, my_pink, my_blue)
my_theme <- theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold")) +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18)) +
  theme(legend.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 16)) +
  theme(strip.text.x = element_text(size = 18, face = "bold"))

dens_plot <- function(fit, pars, include_sim_value = TRUE) {
  plot_data <- data.frame()

  for (p in pars) {
    sim_value <- param_settings[[p]] %||% NA  # if p is not in parameters, sim_value = NA
    dat <- data.frame(
      parameter = rep(p, it),
      estimate = array(fit$draws(p)),
      sim_value = rep(sim_value, it)
    )
    plot_data <- rbind(plot_data, dat)
  }

  plot <- ggplot(plot_data, aes(x = estimate, color = parameter, fill = parameter)) +
    geom_density(alpha = 0.6) +
    labs(title = "Posterior distribution", x = "Estimate", y = "Density", color = "Parameter", fill = "Parameter") +
    scale_color_manual(values = my_colors) +  
    scale_fill_manual(values = my_colors) +
    my_theme
  
  if (include_sim_value) {
    if (any(!is.na(plot_data$sim_value))) {   # only plot sim_value if there's at least one that's not NA
      plot <- plot +
        geom_vline(aes(xintercept = sim_value, color = parameter, linetype = "sim_value")) +
        scale_linetype_manual(values = c("sim_value" = 2), name = NULL)
    }
  }

  return(plot)
}

dot_error_plot <- function(fit, pars, include_sim_value = TRUE) {
  plot_data <- data.frame()

  cred_int <- function(posterior_dist) {
    return(as.numeric(quantile(posterior_dist, c(0.025, 0.975))))
  }

  for (p in pars) {
    dat <- data.frame(
      parameter = p,
      mean = mean(array(fit$draws(p))),
      cred_int_min = cred_int(array(fit$draws(p)))[1],
      cred_int_max = cred_int(array(fit$draws(p)))[2],
      sim_value = param_settings[[p]] %||% NA  # if p is not in parameters, sim_value = NA
    )
    plot_data <- rbind(plot_data, dat)
  }

  plot <- ggplot(plot_data, aes(x = parameter, y = mean, color = parameter)) +
    geom_point() +
    geom_errorbar(aes(ymin = cred_int_min, ymax = cred_int_max), width = 0.25) +
    labs(x = element_blank(), y = "Estimate") +
    scale_color_manual(values = my_colors) +
    guides(color = "none") +
    my_theme

  if (include_sim_value) {
    if (any(!is.na(plot_data$sim_value))) {   # only plot sim_value if there's at least one that's not NA
      plot <- plot +
        geom_hline(aes(yintercept = sim_value, color = parameter, linetype = "sim_value")) +
        scale_linetype_manual(values = c("sim_value" = 2), name = NULL)
    }
  }
  
  return(plot)
}

```
```{r inspect-model, warning = FALSE}
dens_plot(fit, "LR")
dens_plot(fit, "inv_temp")

dot_error_plot(fit, c("LR", "inv_temp"))
```

# Some next steps
1. Fix the estimates;
1. Implement confirmation bias;
1. Create a `generated quantities` block to see what Q values, choices and ratings the model predicts;
1. Hierarchically estimate parameter values for participants;
1. "Cosmetic" improvement: remove the temporary `Q_t` vector;
1. "Cosmetic" improvement in the trial loop. The code for updating values is a bit unwieldy now, so I would prefer it to look something like this if possible:
```
Q[t+1, choice[t]] = Q[t, choice[t]] + LR * pred_error
Q[t+1, !choice[t]] = Q[t, !choice[t]]
```