---
title: "Simulation showcase"
author: "Gilles Lijnzaad"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: TRUE
    number_sections: TRUE
    highlight: tango
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(purl = knitr::hook_purl)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 4)
```

In this document, I demonstrate the behavior of the simulation by varying the parameter settings and visualizing the resulting behavior. The full simulation code can be found in [sim.R](sim.R), and all plot-related functions come from [plot_utils.R](../plot_utils.R).

# Simulation code {.unnumbered}
For ease of reference, here is the simulation code.
```{r sim-code, comment = NA}
rm(list = ls())
sim <- new.env()
source("sim.R", local = sim)  # access functions using sim$fun()
temp <- readLines("sim.R")
start <- grep("# === run_std", temp)
end <- grep("# === end of run_std", temp) - 1
cat(temp[start:end], sep = "\n")
```

# Standard settings
```{r run-std}
plot <- new.env()
source("../plot_utils.R", local = plot)  # access functions using plot$fun()

params_std <- list(
  n_part = 50,
  n_trials = 30,
  LR = 0.4,
  inv_temp = 0.5,
  initQF = 8,
  initQU = 2,
  mu_R = c(5, 5), # F and U
  sigma_R = 2
)

dat <- sim$run_std(params_std)
plot$sim_plots(dat, params_std)
```

## Varying sigma_R and initQ
```{r run-std-sigma-init}
params <- modifyList(params_std, list(sigma_R = 3))
dat <- sim$run_std(params)
plot$sim_plots(dat, params)

params <- modifyList(params_std, list(initQF = 2, initQU = 8))
dat <- sim$run_std(params)
plot$sim_plots(dat, params)
```

## Varying learning rate
```{r run-std-LR}
params <- modifyList(params_std, list(LR = 0.2))
dat <- sim$run_std(params)
plot$sim_plots(dat, params)

params <- modifyList(params_std, list(LR = 0.8))
dat <- sim$run_std(params)
plot$sim_plots(dat, params)
```

## Varying inverse temperature
```{r run-std-inv-temp}
params <- modifyList(params_std, list(inv_temp = 0))
dat <- sim$run_std(params)
plot$sim_plots(dat, params)

params <- modifyList(params_std, list(inv_temp = 1.5))
dat <- sim$run_std(params)
plot$sim_plots(dat, params)
```

# Confirmation bias

The [decision tree](../confirmation_bias_overview/confirmation_bias_models_v3.pdf) shows the many options we have for implementing confirmation bias.

The first differentiation is whether bias occurs at the level of learning (*LRN*) or the level of rating (*RTN*).

## Learning: *LRN*
We assume that the relationship between confirmation and learning rate (LR) is either discrete (discr) or continuous (cont).

### Discrete: *discr*
Confirmatory and disconfirmatory outcomes have separate learning rates, where LR_conf > LR_disconf.

#### Close to belief: *approx* {.tabset}
$$LR =
\begin{cases}
{LR}_\text{conf}      &\text{if}\ \left| R - \text{belief} \right| \leq \text{margin} \\
LR_\text{disconf}   &\text{else}
\end{cases}$$
If the difference between the rating and belief is lesser than or equal to the defined margin, the outcome is treated as confirmatory. Else disconfirmatory. Name `approx` comes from the LaTeX code for the "approximately equal to" sign: $R \approx \text{belief}$.

```{r run-LRN-discr-approx}
params_LRN_discr <- list(
  n_part = 50,
  n_trials = 30,
  LRs = list(conf = 0.8, disconf = 0.2),
  inv_temp = 0.5,
  initQF = 8,
  initQU = 2,
  mu_R = c(5, 5), # F and U
  sigma_R = 2,
  margin = 2
)
```

##### *stat* {.unnumbered}
Beliefs are the static "outside-world values" that one has before starting the experiment.
```{r run-LRN-discr-approx-stat}
dat <- sim$run_LRN(params_LRN_discr, sim$LR_approx, "stat")
plot$sim_plots(dat, params_LRN_discr)
```

##### *dyn* {.unnumbered}
Beliefs are the dynamic "within-experiment values".

```{r run-LRN-discr-approx-dyn}
dat <- sim$run_LRN(params_LRN_discr, sim$LR_approx, "dyn")
plot$sim_plots(dat, params_LRN_discr)
```

#### Greater than or equal to belief: *geq* {.tabset}
$$LR =
\begin{cases}
{LR}_\text{conf}      &\text{if}\ R + \text{margin} \geq \text{belief} \\
LR_\text{disconf}   &\text{else}
\end{cases}$$
If the sum of the rating and the defined margin is equal to or larger than the rating, the outcome is treated as confirmatory. Else disconfirmatory. Name `geq` comes from the LaTeX code for the "greater than or equal to" sign: $R \geq \text{belief}$.


##### *stat* {.unnumbered}
Beliefs are the static "outside-world values" that one has before starting the experiment.
```{r run-LRN-discr-geq-stat}
dat <- sim$run_LRN(params_LRN_discr, sim$LR_geq, "stat")
plot$sim_plots(dat, params_LRN_discr)
```

##### *dyn* {.unnumbered}
Beliefs are the dynamic "within-experiment values".

```{r run-LRN-discr-geq-dyn}
dat <- sim$run_LRN(params_LRN_discr, sim$LR_geq, "dyn")
plot$sim_plots(dat, params_LRN_discr)
```

### Continuous: *cont*

## Rating: *RTN*