---
title: "Stan models"
author: "Gilles Lijnzaad"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: TRUE
    number_sections: TRUE
    highlight: tango
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(purl = knitr::hook_purl) # creates an R file upon knitting the Rmd
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 4)
```

# Helper files {.unnumbered}
In this R markdown I use code from the following files: [sim.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#simR), [plot_utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#plot_utilsR), [utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#utilsR).

# The data
## Running the simulation
For the full simulation code, see [run_std()](../R_simulation/sim_showcase.html#0_Simulation_code).
```{r create-data, comment = NA}
rm(list = ls())
# setwd("~/research/climate-RL-mod/stan_models")
main_dir <- "~/research/climate-RL-mod/"
sim_dir <- paste0(main_dir, "R_simulation/")

sim <- new.env()
source(paste0(sim_dir, "sim.R"), local = sim) # access functions using sim$fun()

params <- list(
  n_part = 10,
  n_trials = 30,
  LR_group = 0.4,
  inv_temp_group = 0.5,
  initQ_group = list(F = 8, U = 2),
  mu_R_group = list(F = 5, U = 5),
  sigma_R_group = 2
)

sim_dat <- sim$run_std(params)

cat(paste0("PARAMETER SETTINGS:"), capture.output(dplyr::glimpse(params)), sep = "\n")
cat(paste0("SIMULATED DATA:"), capture.output(dplyr::glimpse(sim_dat)), sep = "\n")
```

## Inspecting the data
Just so I know what to expect of our model, I include plots of the simulated data: Q values and choices over time.

```{r inspecting-data, warning = FALSE}
plot <- new.env()
source(paste0(main_dir, "plot_utils.R"), local = plot) # access functions using plot$fun()

plot$sim_plots(sim_dat, params)
```

# Explaining the model
I show the Stan model code bit by bit to explain what is happening. The full code can be found in [climate-RL.stan](climate-RL.stan).

## `data` block
The data we supply to the model consist of:

1. the number of participants: `n_part`;
1. the number of trials per participant: `n_trials`;
1. an array of choices that a participant made: `choice[n_part, n_trials]`;
1. an array of ratings that a participant gave: `R[n_part, n_trials]`. These are analogous to the array `outcomes[n_trials]` in the `2arm_bandit_example` (see [Stan code](https://github.com/CCS-Lab/hBayesDM/blob/develop/commons/stan_files/bandit2arm_delta.stan) for this example).

```{r explmod-data, comment = NA}
mod_code <- readLines("climate-RL.stan")
start <- grep("data \\{", mod_code)[1]
end <- grep("transformed data \\{", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `transformed data` block
Currently unused.
```{r explmod-transf-data, eval = FALSE, comment = NA}
start <- grep("transformed data \\{", mod_code)
end <- grep("parameters \\{", mod_code)[1] - 2
cat(mod_code[start:end], sep = "\n")
```

## `parameters` block
In the current model, we have two free parameters:

1. learning rate;
1. inverse temperature;

[explanation here]

```{r explmod-params, comment = NA}
start <- grep("parameters \\{", mod_code)[1]
end <- grep("transformed parameters \\{", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `transformed parameters` block
[explanation here]
```{r explmod-transf-params, eval = FALSE, comment = NA}
start <- grep("transformed parameters \\{", mod_code)
end <- grep("model \\{", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `model` block
### The priors
[explanation here]
```{r explmod-model-priors, comment = NA}
start <- grep("model \\{", mod_code)
end <- grep("participant loop", mod_code)[1] - 2
cat(mod_code[start:end], sep = "\n")
```

### Participant loop
For every participant I initialize an array for holding the Q values. I also create a temporary vector `Q_t` that holds the two Q values (F, U) for the *current* trial, because otherwise the `categorical_logit` wouldn't work, somehow. This works for now, and maybe in the future I will find a more elegant solution.
```{r explmod-model-part, comment = NA}
start <- grep("participant loop", mod_code) + 1
end <- grep("trial loop", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

### Trial loop
```{r explmod-model-trial, comment = NA}
start <- grep("trial loop", mod_code) + 1
end <- grep("generated quantities", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `generated quantities` block
[explanation here]
```{r explmod-gen-quant, comment = NA}
start <- grep("generated quantities", mod_code)
end <- length(mod_code)
cat(mod_code[start:end], sep = "\n")
```

# Fitting the model
## Run it
To save running time, I only refit the model if I changed the model or if the simulated data have changed (i.e., if I changed parameter settings). To this end, I compare the previously saved `sim_data.json` to the current `sim_dat`. After this comparison, I can save the new `sim_data` and refit the model if necessary.

```{r run-model}
data_file <- paste0(sim_dir, "sim_dat.json")
dat_changed <- sim$did_sim_dat_change(data_file, sim_dat)
sim$save_sim_dat(params, sim_dat)
model_changed <- FALSE

library(cmdstanr)
options(mc.cores = parallel::detectCores())
if (dat_changed | model_changed) {
  m <- cmdstan_model("climate-RL.stan")
  it <- 1000
  fit <- m$sample(
    data = data_file,
    iter_sampling = it,
    chains = 1,
    thin = 1,
    iter_warmup = it / 2,
    refresh = it / 5,
    seed = 1234
  )
  fit$save_object(file = "climate-RL_fit.rds")
} else {
  fit <- readRDS(file = "climate-RL_fit.rds")
}
draws <- posterior::as_draws_df(fit$draws()) # df makes it easier to handle
draws <- draws %>%
  dplyr::rename(
    LR_group = `means[1]`,
    inv_temp_group = `means[2]`,
    `initQ_group$F` = `means[3]`,
    `initQ_group$U` = `means[4]`
  )
```

## Inspect results
### Group-level parameters
```{r posterior-plots, fig.height = 8}
to_inspect <- c("LR_group", "inv_temp_group", "initQ_group$F", "initQ_group$U")
plot$posterior_density(draws, to_inspect, params)
```
```{r posterior-table}
util <- new.env()
source(paste0(main_dir, "utils.R"), local = util) # access functions using util$fun()
util$print_posterior_table(draws, params, to_inspect)
```

### Participant-level parameters
```{r sim-vs-fit, fig.height = 8}
params <- rjson::fromJSON(file = paste0(sim_dir, "sim_param_settings.json"))
source(paste0(main_dir, "plot_utils.R"), local = plot)
plot$pp_level_param_fit(draws, c("LR", "inv_temp", "initQF", "initQU"), params)
```