---
title: "Simulation showcase"
author: "Gilles Lijnzaad"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    number_sections: TRUE
    highlight: tango
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(purl = knitr::hook_purl)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 4)
```

In this document, I demonstrate the behavior of the simulation by varying the parameter settings and visualizing the resulting behavior.

# Standard settings
```{r run-std}
rm(list = ls())
# setwd("~/research/climate-RL-mod/simulation_showcase")
main_dir <- "~/research/climate-RL-mod/"
model_dir <- paste0(main_dir, "models/")
util_dir <- paste0(main_dir, "utilities/")

plot <- new.env()
source(paste0(util_dir,"plot_utils.R"), local = plot)  # access functions using plot$fun()

params_std <- list(
  n_part = 50,
  n_trials = 30,
  LR_group = 0.4,
  inv_temp_group = 0.5,
  initQF_group = 8,
  initQU_group = 2,
  mu_R_group = list(F = 5, U = 5),
  sigma_R_group = 2
)

sim <- new.env()
source(paste0(model_dir, "1_std.R"), local = sim)  # access functions using sim$fun()

dat <- sim$run(params_std)
plot$sim_plots(dat, params_std)
```

## Varying learning rate
```{r run-std-LR}
params <- modifyList(params_std, list(LR_group = 0.2))
dat <- sim$run(params)
plot$sim_plots(dat, params)

params <- modifyList(params_std, list(LR_group = 0.8))
dat <- sim$run(params)
plot$sim_plots(dat, params)
```

## Varying inverse temperature
```{r run-std-inv-temp}
params <- modifyList(params_std, list(inv_temp_group = 0))
dat <- sim$run(params)
plot$sim_plots(dat, params)

params <- modifyList(params_std, list(inv_temp_group = 1.5))
dat <- sim$run(params)
plot$sim_plots(dat, params)
```

# Confirmation bias

The [decision tree](../confirmation_bias_overview/confirmation_bias_models_v3.pdf) shows the many options we have for implementing confirmation bias.

The first differentiation is whether bias occurs at the level of learning (*LRN*) or the level of rating (*RTN*).

## Learning: *LRN*
We assume that the relationship between confirmation and learning rate (LR) is either discrete (discr) or continuous (cont).

### Discrete: *discr*
Confirmatory and disconfirmatory outcomes have separate learning rates, where LR_conf > LR_disconf.

#### Close to belief: *approx* {.tabset}
$$LR =
\begin{cases}
{LR}_\text{conf}      &\text{if}\ \left| R - \text{belief} \right| \leq \text{margin} \\
LR_\text{disconf}   &\text{else}
\end{cases}$$
If the difference between the rating and belief is lesser than or equal to the defined margin, the outcome is treated as confirmatory. Else disconfirmatory. Name `approx` comes from the LaTeX code for the "approximately equal to" sign: $R \approx \text{belief}$.

```{r params-discr-approx}
params_LRN_discr <- list(
  n_part = 50,
  n_trials = 30,
  LRs_group = list(conf = 0.8, disconf = 0.2),
  inv_temp_group = 0.5,
  initQ_group = list(F = 8, U = 2),
  mu_R_group = list(F = 5, U = 5),
  sigma_R_group = 2,
  margin_group = 2
)
```

##### *stat* {.unnumbered}
Beliefs are the static "outside-world values" that one has before starting the experiment.
```{r run-LRN-discr-approx-stat}
sim <- new.env()
source(paste0(model_dir, "2_LRN_discr_approx_stat.R"), local = sim)

dat <- sim$run(params_LRN_discr)
plot$sim_plots(dat, params_LRN_discr)
```

##### *dyn* {.unnumbered}
Beliefs are the dynamic "within-experiment values".

```{r run-LRN-discr-approx-dyn}
sim <- new.env()
source(paste0(model_dir, "3_LRN_discr_approx_dyn.R"), local = sim)

dat <- sim$run(params_LRN_discr)
plot$sim_plots(dat, params_LRN_discr)
```

#### Greater than or equal to belief: *geq* {.tabset}
$$LR =
\begin{cases}
{LR}_\text{conf}      &\text{if}\ R + \text{margin} \geq \text{belief} \\
LR_\text{disconf}   &\text{else}
\end{cases}$$
If the sum of the rating and the defined margin is equal to or larger than the rating, the outcome is treated as confirmatory. Else disconfirmatory. Name `geq` comes from the LaTeX code for the "greater than or equal to" sign: $R \geq \text{belief}$.


##### *stat* {.unnumbered}
Beliefs are the static "outside-world values" that one has before starting the experiment.
```{r run-LRN-discr-geq-stat}
sim <- new.env()
source(paste0(model_dir, "4_LRN_discr_geq_stat.R"), local = sim)

dat <- sim$run(params_LRN_discr)
plot$sim_plots(dat, params_LRN_discr)
```

##### *dyn* {.unnumbered}
Beliefs are the dynamic "within-experiment values".

```{r run-LRN-discr-geq-dyn}
sim <- new.env()
source(paste0(model_dir, "5_LRN_discr_geq_dyn.R"), local = sim)

dat <- sim$run(params_LRN_discr)
plot$sim_plots(dat, params_LRN_discr)
```

### Continuous: *cont* {.tabset}
The learning rate relates to the deviation from belief in a continuous way, with limits at 0 and predefined LR_max. The relationship is defined by LR_slope.

\begin{alignat*}{2}
LR = LR_{\max} \cdot & LR'\\
& LR' =
\begin{cases}
1                                       &\text{if}\ R +
\text{margin} \geq \text{belief} \\
LR_\text{slope} \cdot \text{diff} + 1   &\text{else}
\end{cases}\\
\text{with } & LR' \in [0, 1]
\end{alignat*}

See figure below (`LR_slope` = 0.25).
```{r illustrate-rel-LR, fig.width = 5, fig.height = 3}
dummy <- data.frame(
  diff = c(-9, 0, 9),
  LR_prime = c(0, 1, 1)
)
ggplot(dummy) + 
  geom_line(aes(x = diff, y = LR_prime)) +
  geom_vline(aes(xintercept = 0), lty = 2) +
  scale_x_continuous(breaks = seq(-9, 9, 3)) +
  scale_y_continuous(breaks = c(0, 1), labels = c("0", expression(w[LR]))) +
  labs(x = "R - belief", y = "LR") +
  plot$my_theme_classic
```

#### *stat* {.unnumbered}
```{r run-LRN-cont-stat}
sim <- new.env()
source(paste0(util_dir, "sim_utils.R"), local = sim)

params_LRN_cont <- list(
  n_part = 50,
  n_trials = 30,
  w_LR_group = 0.8,
  inv_temp_group = 0.5,
  initQ_group = list(F = 8, U = 2),
  mu_R_group = list(F = 5, U = 5),
  sigma_R_group = 2
)

dat <- sim$run_LRN_cont(params_LRN_cont, "stat")
plot$sim_plots(dat, params_LRN_cont)
```

#### *dyn* {.unnumbered}
```{r run-LRN-cont-dyn}
dat <- sim$run_LRN_cont(params_LRN_cont, "dyn")
plot$sim_plots(dat, params_LRN_cont)
```

## Rating: *RTN*