---
title: "Model recovery"
author: "Gilles Lijnzaad"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
    number_sections: FALSE
    highlight: tango
    code_folding: show
---

> In this document I run simulations for each model and fit these data to each other model. I do this 100 times, each time tallying which model won (fits are compared with the `bridgesampling` package). This results in a confusion matrix at the end.

```{r rmd-setup, include = FALSE}
knitr::knit_hooks$set(purl = knitr::hook_purl) # creates an R file upon knitting the Rmd
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 8)
```

```{r setup}
rm(list = ls())
setwd("~/research/climate-RL-mod/model_recovery")
main_dir <- "~/research/climate-RL-mod/"
recov_dir <- paste0(main_dir, "model_recovery/")
util_dir <- paste0(main_dir, "utilities/")
models_dir <- paste0(main_dir, "models/")

n_runs <- 100

settings <- list(
  n_part = 50,
  n_trials = 30,
  mu_R_group = list(F = 5, U = 5),
  sigma_R_group = 2,
  margin_group = 2
)

fitting <- new.env()
source(paste0(util_dir, "fit_utils.R"), local = fitting)
```

# Part A: Fitting the models
## 0. `null` ("guessing") {.tabset}
### Simulate data
```{r sim-0}
sim_model_name <- "0_null"
sim_path <- paste0(models_dir, sim_model_name, ".R")
sim <- new.env()
source(sim_path, local = sim)

sim_dir <- paste0(recov_dir, "dat/", sim_model_name, "/sim/")
if (!dir.exists(sim_dir)) dir.create(sim_dir)

sim$run_many(settings, sim_dir, n_runs)
```

### 0.0 Fit to `null`
```{r fit-0.0}
fit_model_name <- "0_null"

model_path <- paste0(models_dir, fit_model_name, ".stan")
fit_dir <- paste0(recov_dir, "dat/", sim_model_name, "/", fit_model_name, "/")
if (!dir.exists(fit_dir)) dir.create(fit_dir)

# DOES NOT WORK YET
# fitting$fit_many(sim_dir, model_path, fit_dir, n_runs)
```

### 0.1 Fit to `std`
### 0.2 Fit to `LRN_discr_approx_stat`
### 0.3 Fit to `LRN_discr_approx_dyn`
### 0.4 Fit to `LRN_discr_geq_stat`
### 0.5 Fit to `LRN_discr_geq_dyn`

---

## 1. `std` (no confirmation bias) {.tabset}
### Simulate data
```{r sim-1}
sim_model_name <- "1_std"
sim_path <- paste0(models_dir, sim_model_name, ".R")
sim <- new.env()
source(sim_path, local = sim)

sim_dir <- paste0(recov_dir, "dat/", sim_model_name, "/sim/")
if (!dir.exists(sim_dir)) dir.create(sim_dir)

sim$run_many(settings, sim_dir, n_runs)
```

### 1.0 Fit to `0_null`

### 1.1 Fit to `1_std`
```{r fit-1.1}
fit_model_name <- "1_std"

model_path <- paste0(models_dir, fit_model_name, ".stan")
fit_dir <- paste0(recov_dir, "dat/", sim_model_name, "/", fit_model_name, "/")
if (!dir.exists(fit_dir)) dir.create(fit_dir)

# WOULD TAKE AROUND AN HOUR TO RUN
# fitting$fit_many(sim_dir, model_path, fit_dir, n_runs)
```

### 1.2 Fit to `2_LRN_discr_approx_stat`
```{r fit-1.2}
fit_model_name <- "2_LRN_discr_approx_stat"

model_path <- paste0(models_dir, fit_model_name, ".stan")
fit_dir <- paste0(recov_dir, "dat/", sim_model_name, "/", fit_model_name, "/")
if (!dir.exists(fit_dir)) dir.create(fit_dir)

n_runs <- 2
fitting$fit_many(sim_dir, model_path, fit_dir, n_runs)
```

## 2. `LRN_discr_approx_stat`
### Simulate data
```{r sim-2}
sim_model_name <- "2_LRN_discr_approx_stat"
sim_path <- paste0(models_dir, sim_model_name, ".R")
sim <- new.env()
source(sim_path, local = sim)

sim_dir <- paste0(recov_dir, "dat/", sim_model_name, "/sim/")
if (!dir.exists(sim_dir)) dir.create(sim_dir)

settings[["margin_group"]] <- 2
sim$run_many(settings, sim_dir, n_runs)
```

## 3. `LRN_discr_approx_dyn`
## 4. `LRN_discr_geq_stat`
## 5. `LRN_discr_geq_dyn`

---

# Part B: Comparing fits
```{r}
library(bridgesampling)

fit_1_1_a <- readRDS(paste0(recov_dir, "dat/1_std/1_std/fit_001.rds"))
fit_1_1_b <- readRDS(paste0(recov_dir, "dat/1_std/1_std/fit_002.rds"))
fit_1_2_a <- readRDS(paste0(recov_dir, "dat/1_std/2_LRN_discr_approx_stat/fit_001.rds"))
fit_1_2_b <- readRDS(paste0(recov_dir, "dat/1_std/2_LRN_discr_approx_stat/fit_002.rds"))

mod_1 <- cmdstan_model(paste0(models_dir, "1_std.stan"), force_recompile = TRUE)
bridge_fit_1_1_a <- bridge_sampler(fit_1_1_a, mod_1)
# Error: Model methods cannot be used with a pre-compiled Stan executable, the model must be compiled again
```