---
title: "Model fit showcase"
author: "Gilles Lijnzaad"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
    number_sections: TRUE
    highlight: tango
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::knit_hooks$set(purl = knitr::hook_purl) # creates an R file upon knitting the Rmd
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 8)
```

> In this document, I explain the code for my standard Stan model of the climate-RL task (without confirmation bias) and demonstrate parameter recovery.

# Helper files {.unnumbered}
In this R markdown I use code from the following files: [sim_utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#sim_utilsR), [plot_utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#plot_utilsR), [fit_utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#fit_utilsR), and [utils.R](https://gilleslijnzaad.github.io/climate-RL-modeling/all_utils.html#utilsR).

# Creating the data

--- 

## Running the simulation
```{r create-data, comment = NA}
rm(list = ls())
# setwd("~/research/climate-RL-mod/modeling_showcase")
main_dir <- "~/research/climate-RL-mod/"
util_dir <- paste0(main_dir, "utilities/")
mod_dir <- paste0(main_dir, "models/")
model_path <- paste0(mod_dir, "1_std.stan")

sim <- new.env()
source(paste0(mod_dir, "1_std.R"), local = sim) # access functions using sim$fun()

params <- list(
  n_part = 50,
  n_trials = 30,
  LR_group = 0.4,
  inv_temp_group = 0.5,
  initQF_group = 8,
  initQU_group = 2,
  mu_R_group = list(F = 5, U = 5),
  sigma_R_group = 2
)

sim_dat <- sim$run(params)

cat(paste0("PARAMETER SETTINGS:"), capture.output(dplyr::glimpse(params)), sep = "\n")
cat(paste0("SIMULATED DATA:"), capture.output(dplyr::glimpse(sim_dat)), sep = "\n")
```

## Inspecting the data
Just so I know what to expect of our model, I include plots of the simulated data: Q values and choices over time.

```{r inspecting-data, fig.height = 4}
plot <- new.env()
source(paste0(util_dir, "plot_utils.R"), local = plot) # access functions using plot$fun()

plot$sim_plots(sim_dat, params)
```

# Explaining the model

--- 

I show the Stan model code bit by bit to explain what is happening. The full code can be found in [std.stan](std.stan).

## `data` block
The data we supply to the model consist of:

1. the number of participants: `n_part`;
1. the number of trials per participant: `n_trials`;
1. an array of choices that a participant made: `choice[n_part, n_trials]`;
1. an array of ratings that a participant gave: `R[n_part, n_trials]`. These are analogous to the array `outcomes[n_trials]` in the `2arm_bandit_example` (see [Stan code](https://github.com/CCS-Lab/hBayesDM/blob/develop/commons/stan_files/bandit2arm_delta.stan) for this example).

```{r explmod-data, comment = NA}
mod_code <- readLines(model_path)
start <- grep("data \\{", mod_code)[1]
end <- grep("transformed data \\{", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `transformed data` block
Currently unused.
```{r explmod-transf-data, eval = FALSE, comment = NA}
start <- grep("transformed data \\{", mod_code)
end <- grep("parameters \\{", mod_code)[1] - 2
cat(mod_code[start:end], sep = "\n")
```

## `parameters` block
In the current model, we have four free parameters:

1. learning rate;
1. inverse temperature;
1. initial Q for friendly;
1. initial Q for unfriendly.

[explanation here]

```{r explmod-params, comment = NA}
start <- grep("parameters \\{", mod_code)[1]
end <- grep("transformed parameters \\{", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `transformed parameters` block
[explanation here]
```{r explmod-transf-params, eval = FALSE, comment = NA}
start <- grep("transformed parameters \\{", mod_code)
end <- grep("model \\{", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `model` block
### The priors
[explanation here]
```{r explmod-model-priors, comment = NA}
start <- grep("model \\{", mod_code)
end <- grep("participant loop", mod_code)[1] - 2
cat(mod_code[start:end], sep = "\n")
```

### Participant loop
For every participant I initialize an array for holding the Q values. I also create a temporary vector `Q_t` that holds the two Q values (F, U) for the *current* trial, because otherwise the `categorical_logit` wouldn't work, somehow. This works for now, and maybe in the future I will find a more elegant solution.
```{r explmod-model-part, comment = NA}
start <- grep("participant loop", mod_code) + 1
end <- grep("trial loop", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

### Trial loop
```{r explmod-model-trial, comment = NA}
start <- grep("trial loop", mod_code) + 1
end <- grep("generated quantities", mod_code) - 2
cat(mod_code[start:end], sep = "\n")
```

## `generated quantities` block
[explanation here]
```{r explmod-gen-quant, comment = NA}
start <- grep("generated quantities", mod_code)
end <- length(mod_code)
cat(mod_code[start:end], sep = "\n")
```
# Fitting the model

--- 

## Single run: chosen parameters
### Do the run
To save running time, I only refit the model if I changed the model or if the simulated data have changed (i.e., if I changed parameter settings). To this end, I compare the previously saved `sim_data.json` to the current `sim_dat`. After this comparison, I can save the new `sim_data` and refit the model if necessary.

```{r run-model}
dat_dir <- paste0(main_dir, "modeling_showcase/dat/1_run/")
dat_file <- paste0(dat_dir, "sim_dat_001.json")

sim_utils <- new.env()
source(paste0(util_dir, "sim_utils.R"), local = sim_utils) # access functions using sim_utils$fun()

dat_changed <- sim_utils$did_sim_dat_change(dat_file, sim_dat)
sim_utils$save_sim_dat(params, sim_dat, dat_file)
model_changed <- FALSE

fitting <- new.env()
source(paste0(util_dir, "fit_utils.R"), local = fitting) # access functions using fitting$fun()
if (dat_changed | model_changed) {
  model <- cmdstan_model(model_path)
  draws <- fitting$get_draws(model, dat_file, show_iteration_progress = TRUE)
  saveRDS(draws, paste0(dat_dir, "fit_001.rds"))
} else {
  draws <- readRDS(file = paste0(dat_dir, "fit_001.rds"))
}
```

### Group-level parameters
```{r posterior-plots}
to_inspect <- c("LR_group", "inv_temp_group", "initQF_group", "initQU_group")
plot$posterior_density(draws, to_inspect, params)
```
```{r posterior-table}
util <- new.env()
source(paste0(util_dir, "utils.R"), local = util) # access functions using util$fun()
util$print_posterior_table(draws, params, to_inspect)
```

### Participant-level parameters
```{r sim-vs-fit}
participant_params <- rjson::fromJSON(file = paste0(dat_dir, "sim_param_settings_001.json"))
plot$pp_level_param_fit(draws, c("LR", "inv_temp", "initQF", "initQU"), participant_params)
```

Strange: for inv_temp it seems okay, but for the other parameters the participant-level estimations are very similar to the group-level estimations.

## Many runs: random parameters
To check parameter recovery across a range of parameter values, I run the model a bunch of times with different parameter settings. More specifically, each parameter value is chosen randomly from its whole theoretical range.
```{r many-runs}
n_runs <- 100
free_params <- c("LR_group", "inv_temp_group", "initQF_group", "initQU_group")
dat_dir <- paste0(main_dir, "modeling_showcase/dat/")

# running this bit below takes at least an hour
# fitting$sim_fit_many(params, free_params, model_path, dat_dir, n_runs)
```

```{r inspect-many-runs}
sim_params <- data.frame(k = 1:n_runs)
fit_params <- data.frame(k = 1:n_runs)

dat_dir <- paste0(main_dir, "modeling_showcase/dat/100_runs/")

for (k in 1:n_runs) {
  sim_file <- paste0(dat_dir, "sim_param_settings_", sprintf("%03d", k), ".json")
  sim_dat <- rjson::fromJSON(file = sim_file)

  fit_file <- paste0(dat_dir, "fit_", sprintf("%03d", k), ".rds")
  fit_dat <- readRDS(fit_file)

  # calculate medians
  for (p in free_params) {
    sim_params[[p]][k] <- sim_dat[[p]]
    fit_params[[p]][k] <- median(fit_dat[[p]])
  }
}

plot$many_runs_param_fit(sim_params, fit_params, free_params)
```